Weather covariance matrices as U_hyp for GR50

### 2021-01-26 R/Analysis_v0.6_weather_covariance_Uhyp_greenup.Rmd
Add data-driven results draft figure and start Results for this figure

Function to plot covariance matrices to look at their differences
```{r pvdiv_plot_cov function}
pvdiv_plot_cov <- function(covdf){
  U1 <- cor_phe
  if(isSymmetric(U1)){
        for(i in 1:nrow(U1)){
          for(j in 1:ncol(U1)){
            if(i < j){
              U1[i, j] <- NA
            }
          }
        }
  }
  U1 <- as_tibble(U1, rownames = "rowU", .name_repair = "unique") %>%
          pivot_longer(cols = -.data$rowU, names_to = "colU",
                       values_to = "covar") %>%
          filter(!is.na(.data$covar))
  U1$colU <- factor(U1$colU, levels = colnames(cor_phe))
  U1$rowU <- factor(U1$rowU, levels = colnames(cor_phe))

@@

## 3e. Data-driven mat

@@

Example to modify from a previous script:
```{r}
get_U_by_mass(m)
i=2
names(Ulist)[i]
Ulist <- switchgrassGWAS:::get_Ulist(m)
b1 <- svd(Ulist[[i]])
b1$d[1] / sum(b1$d) # Proportion of variance explained by first eigenvector 11.5
b1$d[2] / sum(b1$d) # get to 60%
b1$d[3] / sum(b1$d) # 
b1$d[4] / sum(b1$d) # 
b1$d[5] / sum(b1$d) #

sum(b1$d[1:2]) / sum(b1$d) # First 5 eigenvectors explain 1% or more variance, and 27.7% of total variance.
for(g in 1:5){
barplot(b1$v[,g]/Ulist[[1]][which.max(abs(Ulist[[1]][,g])),g])
}

ggb1 <- tibble(`First eigenvector (PVE = 100%)` = b1$v[,1]/Ulist[[1]][which.max(abs(Ulist[[1]][,1])),1],
       Phenotype = switchgrassGWAS:::get_colnames(m)) %>%
  mutate(Phenotype = str_remove(Phenotype, "Stand_BhatGulf_")) %>%
  separate(Phenotype, into = c("Function", "Garden"), sep = -3) %>%
  mutate(Function = str_remove(Function, "_$"),
         Garden = str_remove(Garden, "^_"))
ggb1$Garden <- factor(ggb1$Garden, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"))
ggb1_plot <- ggb1 %>%
  ggplot(aes(x = Garden, y = `First eigenvector (PVE = 100%)`)) +
  geom_bar(aes(group = Function, color = Function, fill = Function), stat = "identity", position = "dodge") +
  theme(legend.position = "top", panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(hjust = 1, angle = 45)) +
  geom_hline(yintercept = 0, linetype = 2) +
  scale_color_viridis_d(option = "B", end = 0.7, 
                      direction = -1) +
  scale_fill_viridis_d(option = "B", begin = 0.1, end = 0.7, direction = -1) 
save_plot(paste0("../manuscript/plots/SVD_of_greenup_Gulf_", names(Ulist)[i], ".png"), base_height = 4, plot = ggb1_plot)

ggb2 <- tibble(`First eigenvector (PVE = 21.7%)` = b1$v[,2]/Ulist[[1]][which.max(abs(Ulist[[1]][,2])),2],
       Phenotype = switchgrassGWAS:::get_colnames(m)) %>%
  mutate(Phenotype = str_remove(Phenotype, "Stand_BhatGulf_")) %>%
  separate(Phenotype, into = c("Function", "Garden"), sep = -3) %>%
  mutate(Function = str_remove(Function, "_$"),
         Garden = str_remove(Garden, "^_"))
ggb2$Garden <- factor(ggb1$Garden, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"))
ggb2_plot <- ggb2 %>%
  ggplot(aes(x = Garden, y = `First eigenvector (PVE = 21.7%)`)) +
  geom_bar(aes(group = Function, color = Function, fill = Function), stat = "identity", position = "dodge") +
  theme(legend.position = "top", panel.spacing.x = unit(0.1, 'cm'), axis.text.x = element_text(hjust = 1, angle = 45)) +
  geom_hline(yintercept = 0, linetype = 2) +
  scale_color_viridis_d(option = "B", end = 0.7, 
                      direction = -1) +
  scale_fill_viridis_d(option = "B", begin = 0.1, end = 0.7, direction = -1) 
save_plot(paste0("../manuscript/plots/SVD_of_greenup_Gulf_", names(Ulist)[i], "Eigenvector_2.png"), base_height = 4, plot = ggb2_plot)

>>>>

Function to plot covariance matrices to look at their differences
```{r pvdiv_plot_cov function}
pvdiv_plot_cov <- function(covdf){
  U1 <- covdf
  if(isSymmetric(U1)){
        for(i in 1:nrow(U1)){
          for(j in 1:ncol(U1)){
            if(i < j){
              U1[i, j] <- NA
            }
          }
        }
  }
  U1 <- as_tibble(U1, rownames = "rowU", .name_repair = "unique") %>%
          pivot_longer(cols = -.data$rowU, names_to = "colU",
                       values_to = "covar") %>%
          filter(!is.na(.data$covar))
  U1$colU <- factor(U1$colU, levels = colnames(covdf))
  U1$rowU <- factor(U1$rowU, levels = colnames(covdf))

@@

# -------------------------

## 3e. Data-driven matrices

Run functions on line 743 first - to get posterior weights.
```{r}
pw_gamg %>%
  filter(log10BayesFactor >= 2)
pw_gg %>%
  filter(log10BayesFactor >= 2) %>%
  select(-(Marker:value), -log10BayesFactor) %>%
  colSums()/902 
pw_mg %>%
  filter(log10BayesFactor >= 2)
pw_gamf %>%
  filter(log10BayesFactor >= 2)
pw_gf %>%
  filter(log10BayesFactor >= 2)
pw_mf %>%
  filter(log10BayesFactor >= 2)
```
Gulf greenup sig 902 SNPs: 15.1% ED_PCA_1.9 
24.713% ED_tPCA.9
24.315% ED_tPCA.12
5.722 ED_tPCA.13

@@

ED_tPCA is made up of the 5 ED_PCA_#, each of which explains a different PVE of the tPCA. SNPs can have high pw's for just one of these PCA's, and/or for the tPCA, which is a weighted combination of all 5. While the PCA's typically have simple, garden-based effects, the tPCA is a more complicated weighted combination of these effects.

### i. Function

```{r functions to plot svd of data driven covar matrices}
get_svd_df <- function(m, i = NULL, eigen_v = 1){
  Ulist <- switchgrassGWAS:::get_Ulist(m)
  stopifnot(eigen_v %in% c(1,2,3,4,5))
  if(is.null(i)){
    i_v <- which(names(Ulist) %in% names(Ulist)[get_U_by_mass(m, thresh = 0.1)])
    i = i_v[1]
  } else if(is.character(i)){
    i <- which(names(Ulist) %in% i)
  }
  stopifnot(i %in% 1:length(Ulist))

  b1 <- svd(Ulist[[i]])
  PVE_EV <- b1$d[eigen_v] / sum(b1$d) # Proportion of variance explained by this eigenvector
  ggb1 <- tibble(`Eigenvector` = b1$v[,eigen_v]/Ulist[[i]][which.max(abs(Ulist[[i]][,eigen_v])),eigen_v],
       Phenotype = switchgrassGWAS:::get_colnames(m)) %>%
  mutate(Phenotype = str_remove(Phenotype, "Stand_BhatGulf_and_Midwest_")) %>%
  mutate(Phenotype = str_remove(Phenotype, "Stand_BhatGulf_")) %>%
  mutate(Phenotype = str_remove(Phenotype, "Stand_BhatMidwest_")) %>%
  separate(Phenotype, into = c("Phenotype", "Garden"), sep = -3) %>%
  mutate(Phenotype = str_remove(Phenotype, "_$"),
         Garden = str_remove(Garden, "^_"),
         PVE_EV = PVE_EV)
  ggb1$Garden <- factor(ggb1$Garden, levels = c("TX1", "TX2", "TX3", "OK", "MO", "NE", "MI", "SD"))
  return(ggb1)
}

mash_plot_svd <- function(m, i = NULL, eigen_v = 1){
  b1_df <- get_svd_df(m = m, i = i, eigen_v = eigen_v)
  if(eigen_v == 1){
    ylab <- paste0("First eigenvector (PVE = ", eval(round(b1_df$PVE_EV[1]*100,1)), "%)")
  } else if(eigen_v == 2){
    ylab <- paste0("Second eigenvector (PVE = ", eval(round(b1_df$PVE_EV[1]*100,1)), "%)")
  } else if(eigen_v == 3){
    ylab <- paste0("Third eigenvector (PVE = ", eval(round(b1_df$PVE_EV[1]*100,1)), "%)")
  } else if(eigen_v == 4){
    ylab <- paste0("Fourth eigenvector (PVE = ", eval(round(b1_df$PVE_EV[1]*100,1)), "%)")
  } else if(eigen_v == 5){
    ylab <- paste0("Fifth eigenvector (PVE = ", eval(round(b1_df$PVE_EV[1]*100,1)), "%)")
  }
  ggb1 <- b1_df %>%
    ggplot(aes(x = Garden, y = `Eigenvector`)) +
    switchgrassGWAS::theme_oeco + 
    geom_bar(stat = "identity", position = "dodge") +
    theme(legend.position = "top", panel.spacing.x = unit(0.1, 'cm'),
          axis.text.x = element_text(hjust = 1, angle = 45)) +
    geom_hline(yintercept = 0, linetype = 2) +
    scale_color_viridis_d(option = "B", end = 0.7, 
                        direction = -1) +
    scale_fill_viridis_d(option = "B", begin = 0.1, end = 0.7, 
                         direction = -1)  +
    labs(y = ylab)
  return(ggb1)
}
```

```{r}
gxe_gg <- get_GxE(m_gulf_gr_hyp)
gxe_mg <- get_GxE(m_midw_gr_hyp)
gxe_gamg <- get_GxE(m_gam_gr_hyp)
gxe_gf <- get_GxE(m_gulf_fl_hyp)
gxe_mf <- get_GxE(m_midw_fl_hyp)
gxe_gamf <- get_GxE(m_gam_fl_hyp)
```

```{r}
f3a <- mash_plot_svd(m_gulf_gr_hyp, i = "ED_tPCA") 
f3b <- mash_plot_svd(m_gulf_gr_hyp, i = "ED_tPCA", eigen_v = 2) 
f3e <- mash_plot_svd(m_midw_gr_hyp, i = "ED_tPCA") 
f3f <- mash_plot_svd(m_midw_gr_hyp, i = "ED_tPCA", eigen_v = 2) 
f3i <- mash_plot_svd(m_gam_gr_hyp, i = "ED_tPCA") 
f3j <- mash_plot_svd(m_gam_gr_hyp, i = "ED_tPCA", eigen_v = 2) 
f3m <- mash_plot_svd(m_gulf_fl_hyp, i = "ED_tPCA") 
f3n <- mash_plot_svd(m_gulf_fl_hyp, i = "ED_tPCA", eigen_v = 2) 
f3q <- mash_plot_svd(m_midw_fl_hyp, i = "ED_tPCA") 
f3r <- mash_plot_svd(m_midw_fl_hyp, i = "ED_tPCA", eigen_v = 2) 
f3u <- mash_plot_svd(m_gam_fl_hyp, i = "ED_tPCA") 
f3v <- mash_plot_svd(m_gam_fl_hyp, i = "ED_tPCA", eigen_v = 2)

f3c <- mash_plot_pairwise_sharing(corrmatrix = gxe_gg$S_AP, reorder = FALSE)
f3d <- mash_plot_pairwise_sharing(corrmatrix = gxe_gg$S_DS, reorder = FALSE)
f3g <- mash_plot_pairwise_sharing(corrmatrix = gxe_mg$S_AP, reorder = FALSE)
f3h <- mash_plot_pairwise_sharing(corrmatrix = gxe_mg$S_DS, reorder = FALSE)
f3k <- mash_plot_pairwise_sharing(corrmatrix = gxe_gamg$S_AP, reorder = FALSE)
f3l <- mash_plot_pairwise_sharing(corrmatrix = gxe_gamg$S_DS, reorder = FALSE)
f3o <- mash_plot_pairwise_sharing(corrmatrix = gxe_gf$S_AP, reorder = FALSE)
f3p <- mash_plot_pairwise_sharing(corrmatrix = gxe_gf$S_DS, reorder = FALSE)
f3s <- mash_plot_pairwise_sharing(corrmatrix = gxe_mf$S_AP, reorder = FALSE)
f3t <- mash_plot_pairwise_sharing(corrmatrix = gxe_mf$S_DS, reorder = FALSE)
f3w <- mash_plot_pairwise_sharing(corrmatrix = gxe_gamf$S_AP, reorder = FALSE)
f3x <- mash_plot_pairwise_sharing(corrmatrix = gxe_gamf$S_DS, reorder = FALSE)
```
put labels in latitudinal order
get rid of all the STand_Bhat_pop_GR50_ garbage

```{r}
ggdraw() +
  draw_plot(f3a, x = 0, y = 0.825, width = 0.33, height = 0.175) +
  draw_plot(f3b, x = 0, y = 0.65, width = 0.33, height = 0.175) +
  draw_plot(f3d$gg_corr, x = 0.165, y = 0.5, width = 0.165, height = 0.15, scale = 2) +
  draw_plot(f3c$gg_corr, x = 0, y = 0.5, width = 0.165, height = 0.15, scale = 2) +
  draw_plot(f3e, x = 0.33, y = 0.825, width = 0.33, height = 0.175) +
  draw_plot(f3f, x = 0.33, y = 0.65, width = 0.33, height = 0.175) +
  draw_plot(f3h$gg_corr, x = 0.165+.33, y = 0.5, width = 0.165, height = 0.15, scale = 2) +
  draw_plot(f3g$gg_corr, x = 0+.33, y = 0.5, width = 0.165, height = 0.15, scale = 2)  +
  draw_plot(f3i, x = 0.66, y = 0.825, width = 0.33, height = 0.175) +
  draw_plot(f3j, x = 0.66, y = 0.65, width = 0.33, height = 0.175) +
  draw_plot(f3l$gg_corr, x = 0.165+.66, y = 0.5, width = 0.165, height = 0.15, scale = 2) +
  draw_plot(f3k$gg_corr, x = 0+.66, y = 0.5, width = 0.165, height = 0.15, scale = 2) +  ## greenup above here, flowering below
  draw_plot(f3m, x = 0, y = 0.825-0.5, width = 0.33, height = 0.175) +
  draw_plot(f3n, x = 0, y = 0.65-0.5, width = 0.33, height = 0.175) +
  draw_plot(f3p$gg_corr, x = 0.165, y = 0.5-0.5, width = 0.165, height = 0.15, scale = 2) +
  draw_plot(f3o$gg_corr, x = 0, y = 0.5-0.5, width = 0.165, height = 0.15, scale = 2) +
  draw_plot(f3q, x = 0.33, y = 0.825-0.5, width = 0.33, height = 0.175) +
  draw_plot(f3r, x = 0.33, y = 0.65-0.5, width = 0.33, height = 0.175) +
  draw_plot(f3t$gg_corr, x = 0.165+.33, y = 0.5-0.5, width = 0.165, height = 0.15, scale = 2) +
  draw_plot(f3s$gg_corr, x = 0+.33, y = 0.5-0.5, width = 0.165, height = 0.15, scale = 2)  +
  draw_plot(f3u, x = 0.66, y = 0.825-0.5, width = 0.33, height = 0.175) +
  draw_plot(f3v, x = 0.66, y = 0.65-0.5, width = 0.33, height = 0.175) +
  draw_plot(f3x$gg_corr, x = 0.165+.66, y = 0.5-0.5, width = 0.165, height = 0.15, scale = 2) +
  draw_plot(f3w$gg_corr, x = 0+.66, y = 0.5-0.5, width = 0.165, height = 0.15, scale = 2)

save_plot(filename = "Figure_3_Data_driven_Covariance_matrices.png", plot = last_plot(), base_height = 8, base_width = 9)
```
